---
title: "R Notebook"
author: "Data Table Format"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
      toc: TRUE
      toc_float: TRUE
---

```{r include=FALSE}
library(lattice)
library(ggplot2)
```


# Dotplot groups on y-axis

```{r include=FALSE}
dat = structure(
  list(
    Commodity = structure(c(1L, 1L, 1L, 1L, 2L, 2L, 
2L, 2L, 2L, 2L, 1L, 1L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 2L), 
.Label = c("Coffee", 
"Tea"), class = "factor"),
rank = c(1L, 2L, 3L, 4L, 5L, 6L, 7L, 
8L, 9L, 10L, 1L, 2L, 3L, 4L, 5L, 6L, 7L, 8L, 9L, 10L),
Country = structure(c(10L, 
3L, 1L, 4L, 9L, 8L, 7L, 6L, 2L, 5L, 10L, 1L, 3L, 4L, 9L, 8L, 
7L, 6L, 2L, 5L), 
.Label = c("Brazil", "China", "Colombia", "Global Coffee", 
"Global Tea", "India", "Kenya", "Sri Lanka", "Turkey", "Vietnam"
), class = "factor"), 
year = c(2010L, 2010L, 2010L, 2010L, 2010L, 
2010L, 2010L, 2010L, 2010L, 2010L, 2009L, 2009L, 2009L, 2009L, 
2009L, 2009L, 2009L, 2009L, 2009L, 2009L), 
tonnes = c(0.55, 0.75, 
1.3, 2.6, 0.2, 0.3, 0.4, 1, 1.6, 3.5, 0.6, 0.9, 0.9, 2.4, 0.1, 
0.1, 0.5, 0.9, 1.7, 3.3)), 
.Names = c("Commodity", "rank", "Country", 
"year", "tonnes"), 
class = "data.frame", 
row.names = c(NA, -20L
))
```

```{r}
gg <- ggplot(data=dat, aes(x=tonnes, y=reorder(Country, rank)))
gg <- gg + geom_point(aes(color=Commodity))
gg <- gg + facet_wrap(~Commodity, ncol=1)
gg <- gg + labs(y="", title="Production by country")
gg <- gg + theme_bw()
gg <- gg + theme(legend.position="none")
gg <- gg + theme(panel.grid.major.x=element_blank())
gg <- gg + theme(panel.grid.minor.x=element_blank())
gg <- gg + theme(strip.background=element_blank())

gg
```

# Diverging Bar Charts – Plotting Variance with ggplot2

### Data manipulation for Mtcars
reference: https://hutsons-hacks.info/diverging-bar-charts-plotting-variance-with-ggplot2
```{r}
mtcars_sample = mtcars

mtcars_sample$CarBrand <- rownames(mtcars_sample) # Create new column for car brands and names

# add a Z score calculation
mtcars_sample$mpg_z_score <- round((mtcars_sample$mpg - mean(mtcars_sample$mpg))/sd(mtcars_sample$mpg), digits = 2)

# Creating a cut off (above/below mean)
mtcars_sample$mpg_type <- ifelse(mtcars_sample$mpg_z_score < 0, "below", "above")

# Convert the car Brand into a unique factoe and to sort by the Z score 
mtcars_sample <- mtcars_sample[order(mtcars_sample$mpg_z_score), ] #Ascending sort on Z Score
mtcars_sample$CarBrand <- factor(mtcars_sample$CarBrand, levels = mtcars_sample$CarBrand)
head(mtcars_sample)
```


### Creating the plot
```{r}
ggplot(mtcars_sample, aes(x = CarBrand, y= mpg_z_score, label=mpg_z_score)) +
  geom_bar(stat = 'identity', aes(fill=mpg_type), width = .5) +
  scale_fill_manual(name="Mileage (deviation)",
                    labels = c("Above Average", "Below Average"),
                    values = c("above"="#00ba38", "below"="#0b8fd3")) +
  labs(subtitle="Z score (normalised) mileage for mtcars'",
         title= "Diverging Bar Plot (ggplot2)", 
       caption="Produced by Gary Hutson") +
   coord_flip()+
   theme_bw()
```

### Creating the dot plot (second) 
```{r}
ggplot(mtcars_sample, aes(x=CarBrand, y=mpg_z_score, label=mpg_z_score)) +
  geom_point(stat='identity', aes(col=mpg_type), size=6) +
  scale_color_manual(name="Mileage (deviation)",
                     labels = c("Above Average", "Below Average"),
                     values = c("above"="#00ba38", "below"="#0b8fd3")) +
  geom_text(color="white", size=2) +
  labs(title="Diverging Dot Plot (ggplot2)",
       subtitle="Z score showing Normalised mileage", caption="Produced by Gary Hutson") +
  ylim(-2.5, 2.5) +
  coord_flip() +
  theme_bw()
```

# Bar Charts


reference: https://uc-r.github.io/barcharts
```{r include=FALSE}
#library(xlsx)           # for reading in Excel data
library(dplyr)          # for data manipulation
library(tidyr)          # for data manipulation
library(magrittr)       # for easier syntax in one or two areas
library(gridExtra)      # for generating some comparison plots
library(ggplot2)        # for generating the visualizations
```

```{r}
head(mtcars)
```

### Plot
```{r}
library(ggplot2)
library(gridExtra)
library(dplyr)

avg_mpg <- mtcars %>%
        group_by(cyl, am) %>%
        summarise(mpg = mean(mpg, na.rm = TRUE))

p1 <- ggplot(avg_mpg, aes(factor(cyl), mpg, fill = factor(am))) +
        geom_bar(stat = "identity", position = position_dodge(width = .5)) +
        ggtitle("Overlap of side-by-side bars") +
        theme(legend.position = "none")

p2 <- ggplot(avg_mpg, aes(factor(cyl), mpg, fill = factor(am)))+
  geom_bar(stat = "identity", position = position_dodge(width = 1)) +
  ggtitle("Spacing between side-by-side bars") +
        labs(fill = "AM") +
  theme(legend.position = c(1,1), 
        legend.justification = c(1,1),
        legend.background = element_blank())

grid.arrange(p1, p2, ncol = 3)
```

### Plot 2
```{r}
# calculate percentages of each cyl & am category

proportion <- 
  mtcars %>%
    group_by(cyl, am) %>%
    tally() %>%
    group_by(cyl) %>%
    mutate(pct = n/sum(n))

# create proportional stacked bars
ggplot(proportion, aes(factor(cyl), pct, fill = factor(am, levels = c(1, 0)))) +
        geom_bar(stat = "identity", color = "grey40") +
        scale_fill_manual(values = c("#a1d99b", "#31a354")) +
        labs(fill = "AM")
```

# Plot 3
```{r}
p1 <- ggplot(mtcars, aes(reorder(row.names(mtcars), mpg), mpg)) +
        geom_bar(stat = "identity") +
        coord_flip() +
        geom_text(aes(label = mpg), nudge_y = 2)

p2 <- ggplot(mtcars, aes(reorder(row.names(mtcars), mpg), mpg)) +
        geom_bar(stat = "identity") +
        coord_flip() +
        geom_text(aes(label = mpg), nudge_y = -2, color = "white")

grid.arrange(p1, p2, ncol = 2)
```

### Plot 4
```{r}
cars <-
  mtcars %>%
  mutate(Make = row.names(mtcars),
         ID = ifelse(Make == "Fiat X1-9", TRUE, FALSE))

ggplot(cars, aes(reorder(Make, mpg), mpg, fill = ID)) +
        geom_bar(stat = "identity") +
        coord_flip() +
        scale_fill_manual(values = c("grey90", "dodgerblue")) +
         annotate("text", x = "Fiat X1-9", y = 22, label = "mpg = 27.3", color = "white", size = 3) +
  theme_bw()
```

### plot 5
```{r}
proportion <- proportion %>%
                group_by(cyl) %>%
                mutate(label_y = cumsum(pct))

ggplot(proportion, aes(factor(cyl), pct, fill = factor(am, levels = c(1, 0)))) +
        geom_bar(stat = "identity", color = "grey40") +
        geom_text(aes(label = round(pct, 2), y = label_y), vjust = 1.5, color = "white") +
        scale_fill_manual(values = c("#a1d99b", "#31a354")) +
        labs(fill = "AM")
```

# Waterfall Chart

```{r}
library(ggplot2)
library(waterfalls)

category <- c("Sales", "Services", "Fixed Costs", 
              "Variable Costs", "Taxes")
amount <- c(101000, 52000, -23000, -15000, -10000)
income <- data.frame(category, amount) 


waterfall(income) 
```

```{r}
# Create waterfall chart with total column 

waterfall(income,
          calc_total = TRUE,
          total_axis_text = "Net",
          total_rect_text_color = "black",
          total_rect_color = "goldenrod1") +
  scale_y_continuous(labels = scales::dollar) +
  labs(title = "West Coast Profit and losses",
       subtitle = "Year 2017",
       y = "",
       x = "") +
  theme_minimal()
```

```{r}
WaterfallData <- data.frame(
          labels = c("Prev yr","jan","feb","mar",
                      "apr","may","jun","jul","aug",
                     "sep","oct","nov","dec","prent yr"), 
         value = c(45,2,-18,-25,6,7.6,9.8,-6.8,-10,7.7,8.5,-16,5,15.8),
          logic = c(TRUE,FALSE,FALSE,FALSE,FALSE,FALSE,
                    FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE))

head(WaterfallData)
```

```{r}
WaterfallData$labels <- factor(WaterfallData$labels, levels = WaterfallData$labels)
WaterfallData$end <- cumsum(WaterfallData$value)
WaterfallData$end <- c(head(WaterfallData$end, -1),0)
WaterfallData$start <- c(0,head(WaterfallData$end, -1))
WaterfallData$id <- seq_along(WaterfallData$value)

WaterfallData$logic1 <- ifelse(WaterfallData$logic == "TRUE","TOTAL",
                          ifelse(sign(WaterfallData$value) == 1,"UP","DOWN" ))
head(WaterfallData,2)
```

```{r}
# https://github.com/vivekvix/R-Charts/blob/master/R_Plot/TopGear_RChart/Pie_Chart

ggplot(WaterfallData, aes(value, fill = logic1)) + 
  geom_rect(aes(x = labels,
                xmin = id - 0.35,
                xmax = id + 0.35,
                ymin = end,
                ymax = start)) +
  ggtitle("WaterfallData Plot to showing financial data") +
  geom_text(data = WaterfallData, 
            aes(x = id, y = pmax(start, end)+2, label = value),
            size = 4) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none")
```

https://learnr.wordpress.com/2010/05/10/ggplot2-waterfall-charts/
```{r}
balance <- data.frame(
              desc = c("Starting Cash","Sales",
                       "Refunds", "Payouts", "Court Losses",
                       "Court Wins", "Contracts", "End Cash"),
              amount = c(2000,3400, -1100, -100, -6600, 3800, 1400, 2800))
balance
```

```{r}
balance$desc <- factor(balance$desc, levels = balance$desc)
balance$id <- seq_along(balance$amount)
balance$type <- ifelse(balance$amount > 0, "in", "out")
head(balance,2)
```
```{r}
balance[balance$desc %in% c("Starting Cash", "End Cash"), "type" ] <- "net"

head(balance,2)
```

```{r}
balance$end <- cumsum(balance$amount)
balance$end <- c(head(balance$end,-1),0)
balance$start <- c(0,head(balance$end,-1))

balance <- balance[,c(3,1,4,6,5,2)]
head(balance,2)
```

# Graph detective
reference: https://livefreeordichotomize.com/2020/05/17/graph-detective/
```{r}
library(tidyverse)
library(ggridges)

d <- read_csv("https://github.com/nytimes/covid-19-data/raw/master/us-counties.csv")
head(d,2)
```


```{r}
d <- d %>%
  filter(state == "Georgia",
         county %in% c("Cobb","DeKalb","Fulton","Gwinnett", "Hall")) %>%
         group_by(county) %>%
         mutate(case = c(cases[1], diff(cases)))
  
d %>%
  filter(date >= "2020-04-26", date <= "2020-05-09") %>%
  mutate(
     date = format(date, "%d%b%Y"),
     date = factor(date,
                  levels = c("28Apr2020", "27Apr2020", "29Apr2020",
                             "01May2020", "30Apr2020", "04May2020",
                             "06May2020", "05May2020", "02May2020",
                             "07May2020", "26Apr2020", "03May2020",
                             "08May2020", "09May2020"))) %>%
 group_by(date) %>%
 mutate(rank = rank(-case, ties = "first")) %>%
 ggplot(aes(x = date, y = case, group = rank, fill = county)) +
  geom_col(position = position_dodge()) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  scale_fill_manual("County",
                    values = c("#5854A8", "#46868E", "#958648",
                               "#8F5D37", "#5D98DB"),
                    guide = guide_legend(title.position = "top", title.hjust = 0.5))+
  ggtitle(
    "Top 5 Counties with the Greatest Number of Confirmed COVID-19 Cases",
    subtitle = "The chart below represents the most impacted counties over the past 15 days and the number of cases over time.") +
  theme_classic() +
  theme(legend.position = "top", 
         panel.background = element_rect(fill = "#182F4E"),
         plot.background = element_rect(fill = "#182F4E"),
         axis.line = element_line(color = "white"),
         axis.text = element_text(color = "white"),
         axis.title = element_text(color = "white"),
         legend.background = element_rect(fill = "#182F4E"),
         legend.text = element_text(color = "white"),
         legend.title = element_text(color = "white"),
         title = element_text(color = "white"),
         panel.grid.major.y = element_line(color = "white", size = 0.2, linetype = "dotted" ))
    

  
```

```{r}
 d %>%
  filter(date >= "2020-04-26", date <= "2020-05-09") %>%
  mutate(
    date = format(date, "%d%b%Y"),
    date = factor(date,
                  levels = c("02May2020", "06May2020", "04May2020",
                             "01May2020", "27Apr2020", "29Apr2020",
                             "28Apr2020", "05May2020", "08May2020",
                             "30Apr2020", "09May2020", "03May2020",
                             "07May2020", "26Apr2020"))) %>%
  group_by(date) %>%
  mutate(rank = rank(-case, ties = "first")) %>%
  ggplot(aes(x = date, y = case, group = rank, fill = county)) +
  geom_col(position = position_dodge()) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  scale_fill_manual("County",
                    values = c("#5854A8", "#46868E", "#958648",
                               "#8F5D37", "#5D98DB"),
                    guide =  guide_legend(title.position = "top",
                                          title.hjust = 0.5)) +
  ggtitle(
    "Top 5 Counties with the Greatest Number of Confirmed COVID-19 Cases",
    subtitle = "The chart below represents the most impacted counties over the past 15 days and \n the number of cases over time.") +
  theme_classic() +
  theme(legend.position = "top",
        panel.background = element_rect(fill = "#182F4E"),
        plot.background = element_rect(fill = "#182F4E"),
        axis.line = element_line(color = "white"),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white"),
        legend.background = element_rect(fill = "#182F4E"),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white"),
        title = element_text(color = "white"),
        panel.grid.major.y = element_line(color = "white",
                                          size = 0.25,
                                          linetype = "dotted" ))

```

```{r}
d %>%
  filter(date >= "2020-04-26", date <= "2020-05-09") %>%
  ggplot(aes(x = date, y = case, group = county, fill = county)) +
  geom_col(position = "identity", alpha = 0.75) + 
  scale_x_date(date_labels = "%d%b%Y", date_breaks = "2 days") +
  scale_fill_manual("County",
                    values = c("#5854A8", "#46868E", "#958648",
                               "#8F5D37", "#5D98DB"),
                    guide = guide_legend(title.position = "top",
                                         title.hjust = 0.5)) + 
  ggtitle(
    "Top 5 Counties with the Greatest Number of Confirmed COVID-19 Cases",
    subtitle = "The chart below represents the most impacted counties over the past 15 days and the number of cases over time.") +
  theme_classic() +
  theme(legend.position = "top",
        panel.background = element_rect(fill = "#182F4E"),
        plot.background = element_rect(fill = "#182F4E"),
        axis.line = element_line(color = "white"),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white"),
        legend.background = element_rect(fill = "#182F4E"),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white"),
        title = element_text(color = "white"),
        panel.grid.major.y = element_line(color = "white",
                                          size = 0.25,
                                          linetype = "dotted" ))
```

```{r}
d %>%
  filter(date >= "2020-04-26", date <= "2020-05-09") %>%
  head()
```


```{r}
d %>%
  filter(date >= "2020-04-26", date <= "2020-05-09") %>%
  ggplot(aes(x = date, y = case, group = county, fill = county)) +
  geom_col() + 
  scale_x_date(date_labels = "%d%b%Y",
               date_breaks = "5 days") +
  scale_fill_manual("County",
                    values = c("#5854A8", "#46868E", "#958648", 
                               "#8F5D37", "#5D98DB"),
                    guide = guide_legend(title.position = "top",
                                         title.hjust = 0.5)) + 
  ggtitle(
    "Top 5 Counties with the Greatest Number of Confirmed COVID-19 Cases",
    subtitle = "The chart below represents the most impacted counties over the past 15 days and the number of cases over time.") +
  theme_classic() +
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "#182F4E"),
        plot.background = element_rect(fill = "#182F4E"),
        axis.line = element_line(color = "white"),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white"),
        legend.background = element_rect(fill = "#182F4E"),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white"),
        title = element_text(color = "white")) +
  facet_grid(county~.)


```



```{r}
d %>%
  filter(date >= "2020-04-26", date <= "2020-05-09") %>%
  ggplot(aes(x = date, y = county, fill = county, height = case)) +
  geom_density_ridges(scale = 5, stat = "identity") + 
  scale_x_date(date_labels = "%d%b%Y",
               date_breaks = "5 days") +
  scale_fill_manual("County",
                    values = c("#5854A8", "#46868E", "#958648",
                               "#8F5D37", "#5D98DB"),
                    guide = guide_legend(title.position = "top",
                                         title.hjust = 0.5)) + 
  ggtitle(
    "Top 5 Counties with the Greatest Number of Confirmed COVID-19 Cases",
    subtitle = "The chart below represents the most impacted counties over the past 15 days and the number of cases over time.") +
  theme_classic() +
   theme(legend.position = "top", 
        panel.background = element_rect(fill = "#182F4E"),
        plot.background = element_rect(fill = "#182F4E"),
        axis.line = element_line(color = "white"),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white"),
        legend.background = element_rect(fill = "#182F4E"),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white"),
        title = element_text(color = "white")) 
```

### LIne Plot
```{r}
d %>%
  filter(date >= "2020-04-26", date <= "2020-05-09") %>%
  ggplot(aes(x = date, y = case, color = county)) +
  geom_line() + 
  geom_point() +
  scale_x_date(date_labels = "%d%b%Y",
               date_breaks = "5 days") +
  scale_color_manual("County",
                     values = c("#5854A8", "#46868E", "#958648",
                                "#8F5D37", "#5D98DB"),
                     guide = guide_legend(title.position = "top",
                                          title.hjust = 0.5)) + 
  ggtitle(
    "Top 5 Counties with the Greatest Number of Confirmed COVID-19 Cases",
    subtitle = "The chart below represents the most impacted counties over the past 15 days and the number of cases over time.") +
  theme_classic() +
  theme(legend.position = "top",
        panel.background = element_rect(fill = "#182F4E"),
        plot.background = element_rect(fill = "#182F4E"),
        axis.line = element_line(color = "white"),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white"),
        legend.background = element_rect(fill = "#182F4E"),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white"),
        title = element_text(color = "white"),
        panel.grid.major.y = element_line(color = "white",
                                          size = 0.25,
                                          linetype = "dotted" ))
```

```{r}
library(plotly)

d %>%
  filter(date >= "2020-04-26", date <= "2020-05-09") %>%
  ggplot(aes(x = date, y = case, fill = county)) +
  geom_col(position = "identity", alpha = 0.75) + 
  scale_x_date(date_labels = "%d%b%Y",
               date_breaks = "5 days") +
  scale_fill_manual("County",
                    values = c("#5854A8", "#46868E", "#958648",
                               "#8F5D37", "#5D98DB"),
                    guide = guide_legend(title.position = "top",
                                         title.hjust = 0.5)) + 
  ggtitle(
    "Top 5 Counties with the Greatest Number of Confirmed COVID-19 Cases",
    subtitle = "The chart below represents the most impacted counties over the past 15 days and the number of cases over time.") +
  theme_classic() +
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "#182F4E"),
        plot.background = element_rect(fill = "#182F4E"),
        axis.line = element_line(color = "white"),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white"),
        legend.background = element_rect(fill = "#182F4E"),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white"),
        title = element_text(color = "white")) -> p

ggplotly(p)
```


# Grouped Bar Chart Plotly
https://stackoverflow.com/questions/59350468/grouped-bar-chart-plotly
```{r}

library(plotly)
library(tidyverse)

data <- data.frame(Year = 2013:2019, 
                   No_Qualification = c(561,575,579, 627, 682, 674, 707), 
                   Lower_secondary_school_qualification = c(678, 690, 686, 759, 710, 770, 798), 
                   Upper_secondary_school_qualification = c(626, 605, 645, 688, 722, 761, 805), 
                   Level_1_3_post_school_certificate = c(631, 651, 657, 660, 672, 703, 762),
                   Level_4_6_certificate_or_diploma = c(753, 780, 763, 848, 844, 815, 863), 
                   Bachelors_degree = c(980,1006,1028, 1077, 1091, 1125, 1151), 
                   Postgraduate_qualification = c(1178, 1245, 1244, 1331, 1399, 1385, 1474))

head(data, 2)
```

```{r}
data %>%
  plot_ly(x = ~Year, y = ~No_Qualification, type = 'bar', 
          name = 'No_Qualification') %>%
  add_trace(y = ~Lower_secondary_school_qualification, 
              name = 'Lower_secondary_school_qualification') %>%
  add_trace(y = ~Upper_secondary_school_qualification,
              name = 'Upper_secondary_school_qualification') %>%
  add_trace(y = ~Level_1_3_post_school_certificate, 
              name = 'Level_1_3_post_school_certificate') %>%
  add_trace(y = ~Level_4_6_certificate_or_diploma, 
              name = 'Level_4_6_certificate_or_diploma') %>%
  add_trace(y = ~Bachelors_degree, 
              name = 'Bachelors_degree') %>%
  add_trace(y = ~Postgraduate_qualification,
              name = 'Postgraduate_qualification') %>%
  layout(yaxis = list(title = 'Count'), barmode = 'group')
```

# Plotly in R: How to make ggplot2 charts interactive with ggplotly
https://www.musgraveanalytics.com/blog/2018/8/24/how-to-make-ggplot2-charts-interactive-with-plotly
```{r}
library(scales)

df <- data.frame(Date= c(2001, 2001, 2002, 2002, 2003,2003,2004,2004), 
                   Revenue = c(980,1006,1028, 1077, 1091, 1125, 1151,2000), 
                   Target =  c(1178, 1245, 1244, 1331, 1399, 1385, 1474, 1500))

p <- ggplot(data = df,
            aes(x = Date, 
                y = Revenue/100,
                group = 1, 
                text = paste0("Date: ", Date,
                              "<br>Revenue: $", round(Revenue/100, digits = 1), "million",
                              "<br>Target: $", round(Target/100, digits = 1), "million")))+
  scale_y_continuous(labels= scales::dollar_format(prefix="$",suffix=" m"))+
  geom_line(colour = "grey", aes(Date, Target/100)) +
  geom_line(colour = "#408FA6") +
  ylab("Revenue")

ggplotly(p, tooltip = "text")
```

https://stackoverflow.com/questions/34573791/format-hover-text-to-in-stacked-area-plot-r-plot-ly
```{r}
library(plotly)

season <- c('Winter', 'Spring', 'Summer', 'Fall')
y1_real<- c(40, 80, 95, 10)
y2_real<- c(20, 55, 10, 10)
y3_real<- c(40, 10, 60, 180)

df= data.frame(season, y1_real, y2_real, y3_real)

df$y1_reald <- paste(y1_real, "$", sep="")
df$y2_reald <- paste(y2_real, "$", sep="")
df$y3_reald <- paste(y3_real, "$", sep="")

p1<- plot_ly(df, x=season, y=y1_real, text=df$y1_reald, hoverinfo='x+text+name', name="Blue", fill="tonexty")
p2<- add_trace(p1, x=season, y=y2_real, text=df$y2_reald, hoverinfo='x+text+name', name="Orange", fill="tonexty")
p3<- add_trace(p2, x=season, y=y3_real, text=df$y3_reald, hoverinfo='x+text+name', name="Green", fill="tonexty")

p3

#p4<- layout(yaxis=list(showgrid=FALSE, showline=FALSE, hoverformat="$,f"))
#p4 
```



# Style Graphs in ggplot2
http://austinclemens.com/blog/2014/07/03/fivethirtyeight-com-style-graphs-in-ggplot2/
```{r}
library(ggplot2)
library(grid)

line1 = rnorm(100, mean = 15-seq(1, 6, by = .05), sd = 1)
line2 = rnorm(100, mean = 4+seq(1, 21, by=.2), sd=.5)
time = seq(1,100, by = 1)
data = data.frame(time, line1, line2)

head(data, 2)
```

```{r}
ggplot(data, aes(time)) +
  # The actual line
  geom_line(aes(y=line1), size = 1.6 ,colour="#f8766d") +
  geom_line(aes(y = line2), size = 1.6, colour="#00bdc4") +
  theme_bw() +
  # Set the entire chart region to a light gray col
  theme(panel.background = element_rect("#F0F0F0")) +
  theme(plot.background = element_rect("#F0F0F0")) +
  theme(panel.border = element_rect(colour = '#F0F0F0')) +
  # format the grid
  theme(panel.grid.major = element_line(colour = '#D0D0D0', size = .75)) +
  scale_x_continuous(minor_breaks = 0, breaks = seq(0,100,10)) +
  scale_y_continuous(minor_breaks = 0, breaks = seq(0,26,4), limits = c(0,26)) +
  theme(axis.ticks = element_blank()) +
  # Dispose of the legend
  theme(legend.position = "none") +
  # Set title and axis labels, and format these and tick marks
  ggtitle("Some Random Data I Made") +
  theme(plot.title=element_text(face="bold",hjust=-.08,vjust=1,colour="#3C3C3C",size=20)) +
  ylab("Data Label") +
  xlab("Days Since Beginning") +
  theme(axis.text.x=element_text(size=11,colour="#535353",face="bold")) +
  theme(axis.text.y=element_text(size=11,colour="#535353",face="bold")) +
  theme(axis.title.y=element_text(size=11,colour="#535353",face="bold",vjust=1.5)) +
  theme(axis.title.x=element_text(size=11,colour="#535353",face="bold",vjust=-.5)) +
  # Big bold line at y=0
  geom_hline(yintercept=0,size=1.2,colour="#535353") +
  # Plot margins and finally line annotations
  theme(plot.margin = unit(c(1, 1, .5, .7), "cm")) +
  annotate("text",x=14.7,y=16.3,label="Line 1",colour="#f8766d") +
  annotate("text",x=25,y=7.5,label="Line 2",colour="#00bdc4")
 
#ggsave(filename="/538.png",dpi=600)

```

# ANOTHER WAY TO PRESENT LINE GRAPH
reference: https://jiwanheo.rbind.io/blog/2020-01-11-another-way-to-present-line-graph/
```{r}
library(tidyverse)

txhousing %>% head(2)
```

```{r}
txhousing %>%
  ggplot(aes(date, median, color = city)) +
  geom_line() +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_light() +
  theme(legend.position = "none") +
  labs(title = "How did housing prices of Texas change over the years?",
       x = "Year", 
       y = "Median price") 

```

```{r}
tx <- highlight_key(txhousing, ~city)
plot_ly(tx, color = ~city) %>% 
  group_by(city) %>% 
  add_lines(x = ~date, y = ~median) %>% 
  layout(title = "How did housing prices of Texas change over the years?",
         subtitle = "Double-click me!",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Median price"))
```

`highligh_key` declars `city` variables as a "look up" values. What thjis means is you can look up a city in the whole dataset, which effectively isolates the chosen city.

```{r}
library(data.table)

tx_df <- as.data.table(txhousing)

start <- tx_df[tx_df[, .I[date == min(date)], by = city]$V1]
end <- tx_df[tx_df[, .I[date == max(date)], by = city]$V1]

# now we'll stack them and only select data we need 
stacked <- start %>%
              rbind(end) %>%
              select(city, median, date) %>%
              arrange(city)


stacked <- 
      stacked %>%
         spread(date , median) %>%
         rename(start  = `2000`, end = `2015.5`) %>%
         filter(!is.na(start) & !is.na(end))

stacked %>% 
  head()

```

```{r}

stacked <- 
  stacked %>%
    mutate(avg = (start + end)/2) %>%
    mutate(city = fct_reorder(city, avg))          

stacked %>%
  ggplot() +
  geom_segment(aes(x = city, xend = city, y = start, yend = end),
               arrow = arrow(type = "closed", length = unit(0.12, "inches"))) +
  geom_point(aes(city, start), colour = "red", size = 4) +
  geom_point(aes(city, end), colour = "green", size = 4) +
  geom_hline(yintercept = as.numeric(stacked %>% summarize(mean(start))), color = "red", linetype = "dashed") +
  geom_hline(yintercept = as.numeric(stacked %>% summarize(mean(end))), color = "green", linetype = "dashed") +
  geom_text(data = stacked %>% filter(city == "Irving"),
            aes(city, start, label = "Year 2000", hjust = 1.2)) +
   geom_text(data = stacked %>% filter(city == "Irving"),
            aes(city, end, label = "Year 2015", hjust = -0.2)) +
  scale_y_continuous(labels = scales::dollar_format()) +
  coord_flip() +
  labs(title = "How did housing prices of Texas change over the years?",
       x = "City",
       y = "Median price") +
  theme_bw()
```

```{r}
ggplotly(stacked %>% 
  ggplot() +
  geom_segment(aes(x = city, xend = city, y = start, yend = end),
               arrow=arrow(type = "closed", length = unit(0.15, "inches"))) +
  geom_point(aes(city, start), color = "red", size = 4) +
  geom_point(aes(city, end), color = "green", size = 4) +
  geom_hline(yintercept = as.numeric(stacked %>% summarize(mean(start))), color = "red", linetype = "dashed") +
  geom_hline(yintercept = as.numeric(stacked %>% summarize(mean(end))), color = "green", linetype = "dashed") +
  geom_text(data = stacked %>% filter(city == "Irving"),
            aes(city, start, label = "Year 2000", hjust = 0)) +
  geom_text(data = stacked %>% filter(city == "Irving"),
            aes(city, end, label = "Year 2015", hjust = 0)) +
  scale_y_continuous(labels = scales::dollar_format()) +
  coord_flip() +
  labs(title = "How did housing prices of Texas change over the years?",
       x = "City",
       y = "Median price") +
    theme_light())
```

# Shuttering Pies With Retiring Stores
reference: https://rud.is/b/2017/04/21/shuttering-pies-with-retiring-stores/

```{r}
library(hrbrthemes)
library(tidyverse)
library(ggplot2)

data = read.table(text='store,closing,total
"Radio Shack",550,1500
"Payless",400,2600
"Rue21",400,1100
"The Limited",250,250
"bebe",180,180
"Wet Seal",170,170
"Crocs",160,560
"JCPenny",138,1000
"American Apparel",110,110
"Kmart",109,735
"hhgregg",88,220
"Sears",41,695', sep=",", header=TRUE, stringsAsFactors=FALSE)

head(data,3)
```

```{r}
closing_df <-
  data %>%
    as_tibble() %>%
    mutate(remaining = total - closing,
           gone = round((closing/total)*100)/100,
           stay = 1-gone,
           rem_lab = ifelse(remaining == 0, "", scales::comma(remaining))) %>%
    arrange(desc(stay)) %>%
    mutate(store = factor(store, levels = store))

head(closing_df,3)
```

```{r}
ggplot(closing_df) +
  geom_segment(aes(0, store, xend = gone, yend = store, color = "closing"), size = 8) +
  geom_segment(aes(gone, store, xend = gone + stay ,
                   yend = store, color = "Remaining"), size = 8) +
  geom_text(aes(x = 0, y = store, label = closing), color = "white", hjust = 0, nudge_x = 0.01) +
  geom_text(aes(x = 1, y = store, label = rem_lab), color = "white", hjust = 1.5, nudge_x = 0.012) +
  #  scale_x_percent()
  # scale_color_ip
  # scale_color_ipsum(name=NULL) +
  labs(x=NULL, y=NULL, 
       title="Selected 2017 Store closings (estimated)",
       subtitle="Smaller specialty chains such as Bebe and American Apparel are closing their stores,\nwhile lareger chains such as J.C. Penny and Sears are scaling back their footprint.\n") +
  #theme_ipsum_rc(grid="X") +
  theme(axis.text.x=element_text(hjust=c(0, 0.5, 0.5, 0.5, 1))) +
  theme(legend.position=c(0.875, 1.025)) +
  theme(legend.direction="horizontal")
```

# Creating beautiful demographic maps in R with the tidycensus and tmap packages

reference: http://zevross.com/blog/2018/10/02/creating-beautiful-demographic-maps-in-r-with-the-tidycensus-and-tmap-packages/#part-1-using-tidycensus-to-collect-us-census-data


# (ggplot2) Exercising with (ggalt) dumbbells

reference: https://rud.is/b/2016/04/17/ggplot2-exercising-with-ggalt-dumbbells/

```{r}

library(ggplot2) # devtools::install_github("hadley/ggplot2")
library(ggalt)   # devtools::install_github("hrbrmstr/ggalt")
library(dplyr)   # for data_frame() & arrange()

# I'm not crazy enough to input all the data; this will have to do for the example
df <- data_frame(country=c("Germany", "France", "Vietnam", "Japan", "Poland", "Lebanon",
                           "Australia", "South\nKorea", "Canada", "Spain", "Italy", "Peru",
                           "U.S.", "UK", "Mexico", "Chile", "China", "India"),
                 ages_35=c(0.39, 0.42, 0.49, 0.43, 0.51, 0.57,
                           0.60, 0.45, 0.65, 0.57, 0.57, 0.65,
                           0.63, 0.59, 0.67, 0.75, 0.52, 0.48),
                 ages_18_to_34=c(0.81, 0.83, 0.86, 0.78, 0.86, 0.90,
                                 0.91, 0.75, 0.93, 0.85, 0.83, 0.91,
                                 0.89, 0.84, 0.90, 0.96, 0.73, 0.69),
                 diff=sprintf("+%d", as.integer((ages_18_to_34-ages_35)*100)))

head(df)
```

```{r}
# we want to keep the prder in the plot, so we use a factor for country
df <- arrange(df, desc(diff))
df$country <- factor(df$country, levels = rev(df$country))

# We only want first line values with "%" symbols (to avoid chart junk)
# Quick back; there is a more efficient way to do this

head(df)
```

```{r}
# we only want the first line values with '$' symbols (to avoid chart junk)
# quick hack, there is a more efficient way to do this
percent_first <- function(x){
   x <- sprintf("%d%%", round(x*100))
   x[2:length(x)] <- sub("%$","", x[2:length(x)])
   x
}
```

```{r}
gg <- ggplot()
# doing this vs y axis major grid line
gg <- gg + geom_segment(data = df, aes(y=country,yend=country, x=0, xend=1),color="#b2b2b2", size=0.15)

# dumbbell visualization
gg <- gg + geom_dumbbell(data=df, aes(y=country, x=ages_35, xend=ages_18_to_34),
                         size=1.5, color="#b2b2b2", point.size.l=3, point.size.r=3,
                         point.colour.l="#9fb059", point.colour.r="#edae52")

# text below points
gg <- gg + geom_text(data=filter(df, country=="Germany"),
                     aes(x=ages_35, y=country, label="Ages 35+"),
                     color="#9fb059", size=3, vjust=-2, fontface="bold", family="Calibri")

gg <- gg + geom_text(data=filter(df, country=="Germany"),
                     aes(x=ages_18_to_34, y=country, label="Ages 18-34"),
                     color="#edae52", size=3, vjust=-2, fontface="bold", family="Calibri")

# text above points
gg <- gg + geom_text(data=df, aes(x=ages_35, y=country, label=percent_first(ages_35)),
                     color="#9fb059", size=2.75, vjust=2.5, family="Calibri")
gg <- gg + geom_text(data=df, color="#edae52", size=2.75, vjust=2.5, family="Calibri",
                     aes(x=ages_18_to_34, y=country, label=percent_first(ages_18_to_34)))

# difference column
#gg <- gg + geom_rect(data=df, aes(xmin=1.05, xmax=1.175, ymin=-Inf, ymax=Inf), fill="#efefe3")
gg <- gg + geom_text(data=df, aes(label=diff, y=country, x=1.1125), fontface="bold", size=3, family="Calibri")
gg <- gg + geom_text(data=filter(df, country=="Germany"), aes(x=1.1125, y=country, label="DIFF"),
                     color="#7a7d7e", size=3.1, vjust=-2, fontface="bold", family="Calibri")
gg <- gg + scale_x_continuous(expand=c(0,0), limits=c(0, 1.175))
gg <- gg + scale_y_discrete(expand=c(0.075,0))
gg <- gg + labs(x=NULL, y=NULL, title="The social media age gap",
                subtitle="Adult internet users or reported smartphone owners who\nuse social networking sites",
                caption="Source: Pew Research Center, Spring 2015 Global Attitudes Survey. Q74")

# gg <- gg + theme_bw(base_family="Calibri")
# gg <- gg + theme(panel.grid.major=element_blank())
# gg <- gg + theme(panel.grid.minor=element_blank())
# gg <- gg + theme(panel.border=element_blank())
# gg <- gg + theme(axis.ticks=element_blank())
# gg <- gg + theme(axis.text.x=element_blank())
# gg <- gg + theme(plot.title=element_text(face="bold"))
# gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
# gg <- gg + theme(plot.caption=element_text(size=7, margin=margin(t=12), color="#7a7d7e"))

gg
```






