---
title: "Gapminder Data Visaulization"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

The Graphics Companion provides the R code for different data visualisations that can be created using the ggplot2 package.

The Companion adopts the structure of the Financial Times’ Visual Vocabulary by categorising different chart types by the data relationships that they best illustrate.

The data used throughout the Companion derive from a subset of Hans Rosling’s Gapminder World which are available in the gapminder R package. Data on life expectancy at birth, GDP per capita and total population are provided for 142 countries between 1952 and 2007.

LINK: https://www.trafforddatalab.io/graphics_companion/#

# Setup

You need to install - but only once - the tidyverse package and load it into your R session. ggplot2 is part of the tidyverse suite of R tools for data science.
```{r, include=FALSE}
library(tidyverse)
library(gapminder)
library(lubridate)
library(ggplot2)

theme_lab <- function () { 
  theme_grey(base_size = 14, base_family = "sans") #%+replace% 
     theme(
        # plot margin
        plot.margin = unit(rep(0.5, 4), "cm") ,

      # plot background and border
      plot.background = element_blank(),
      panel.background = element_blank(),
      panel.border = element_blank(),

      # grid lines
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(size = 0.5, color = "#cbcbcb", linetype = "dotted"),
      panel.grid.minor = element_blank(),
       
      # axis ticks and lines
      axis.ticks = element_blank(),
      axis.line = element_blank(),

      # title, subtitle and caption
      plot.title = element_text(size = 20, face = "bold", colour = "#757575", hjust = 0),
      plot.subtitle = element_text(size = 16, colour = "#757575", hjust = 0,margin = margin(9, 0, 9, 0)),
      plot.caption = element_text(size = 10, color = "grey50", hjust = 1, margin = margin(t = 15)),

      # axes titles
      axis.title = element_text(colour = "#757575", hjust = 1),
      axis.text.x = element_text(margin = margin(b = 7)),
      axis.text.y = element_text( margin = margin(l = 7)),

      # legend
      legend.position = "top",
      legend.background = element_blank(),
      legend.key = element_blank(),
      legend.title = element_text(size = 12, colour = "#757575"),
      legend.text.align = 0,
      legend.text = element_text(size = 14, colour = "#757575"),

      # facetting
      strip.background = element_rect(fill = "transparent", colour = NA),
      strip.text = element_text(size = 12, face = "bold", colour = "#757575", hjust = 0)
    )
}
```

# Change over time 

### Single line chart

```{r cars}
df <- gapminder %>%
        filter(country == "Argentina") %>%
        mutate(year = as.Date(paste(year,"-01-01", sep = "", format='%Y-%b-%d')))

df %>% 
  ggplot(aes(x=year, y=lifeExp)) +
    geom_line(color = '#fc6721', size = 1) +
    geom_point(color='#fc6721', size = 2) +
    scale_x_date(breaks = df$year, date_labels = "%Y") +
    scale_y_continuous(limits = c(0,max(df$lifeExp)), labels = scales::comma) +
    labs(title = "",
         subtitle = "Life expectancy in Argentina, 1952-2007",
         caption = "Source: Gapminder.org  |  @traffordDataLab",
         x = "",
         y = "Age (years)") +
    theme_lab() +
  theme(panel.grid.major.x = element_blank())
```




### Multiple line chart



```{r pressure, echo=FALSE}
df <- gapminder %>%
        filter(country %in% c("Argentina", "Italy")) %>% 
        mutate(year = as.Date(paste(year, "-01-01", 
                                    sep = "",
                                    format='%Y-%b-%d')))

df %>%
  ggplot(aes(x=year, y=lifeExp, colour=country)) +
  geom_line(size = 1) +
  geom_point(size = 2) + 
  scale_colour_manual(values = c("Argentina" = "#fc6721", "Italy" = "#E7B800")) + 
  scale_x_date(breaks = df$year, date_labels = "%Y") +
   scale_y_continuous(limits = c(0, NA), labels = scales::comma) +
  labs(title = "",
       subtitle = "Life expectancy in Argentina and Italy, 1952-2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = "",
       y = "Age (years)",
       colour = NULL) +
      theme_lab() +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "bottom")
```

### Bar chart (Vertical)
```{r}


df <- 
  filter(gapminder, 
         country == "Egypt") %>% 
  mutate(year = as.Date(paste(year, "-01-01", 
                              sep = "", 
                              format='%Y-%b-%d')))

month_1 = format(as.Date(df$year), "%B")
month_2 = format(as.Date(df$year), "%B")

ggplot(df, aes(x = year, y = gdpPercap)) +
  geom_col(fill = "#fc6721", alpha = 0.8) +
   scale_x_date(breaks = df$year, date_labels = "%Y") +
   scale_y_continuous(expand = c(0, 0), labels = scales::dollar) +
   labs(title = "", 
       subtitle = paste0("GDP per capita in Egypt: ", 
                         month_1,",",min(year(df$year))," - ",
                         month_2,",",max(year(df$year))),
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = NULL, 
       y = "GDP per capita ($)", 
       fill = NULL) +
  theme_lab() + 
  theme(panel.grid.major.x = element_blank())
```

### Slope Chart
```{r}
df <- 
  filter(gapminder,
         country %in% c("Iceland", "Norway") & 
         year %in% c(1952, 2007))

ggplot(df) + 
  geom_line(aes(x = as.factor(year), 
                y = gdpPercap, 
                group = country, 
                colour = country), 
            size = 2, 
            alpha = 0.8) +
  geom_point(aes(x = as.factor(year),
                 y = gdpPercap, 
                 group = country, 
                 colour = country), 
             size = 5, 
             alpha = 0.8) +
  geom_text(data = subset(df, year == 1952), 
            aes(x = as.factor(year), 
                y = gdpPercap, 
                colour = country, 
                label = paste(country, 
                              scales::dollar(
                                        round(gdpPercap, 0)),
                              sep = ", "), 
            size = 4, hjust = 1.2)) +
  geom_text(data = subset(df,
                          year == 2007), 
            aes(x = as.factor(year),
                y = gdpPercap, 
                colour = country, 
                label = scales::dollar(round(gdpPercap, 0))), 
                size = 4, 
                hjust = -0.3
            ) +
  scale_colour_brewer(palette = "Set2") +
  labs(title = "",
       subtitle = "GDP per capita change, 1952 - 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = NULL,
       y = NULL,
       colour = NULL) +
   theme_lab() +
  theme(panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "none") 
  
```

### Stacked area chart

```{r}
df <- gapminder %>% 
  filter(country %in% c("France", "Germany", "Ireland", "Italy")) %>% 
  mutate(year = as.Date(paste(year, "-01-01",
                              sep = "", 
                              format='%Y-%b-%d')))

ggplot(df, aes(x = year, y = gdpPercap, fill = country)) +
  geom_area(color = "white", alpha = 0.4) +
  scale_fill_brewer(palette = "Set2") +
  scale_x_date(breaks = df$year, date_labels = "%Y") +
  scale_y_continuous(expand = c(0, 0), labels = scales::dollar) +
  labs(title = "",
       subtitle = "GDP per capita by country, 1952-2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = NULL,
       y = "GDP per capita ($)",
       fill = NULL) +
  theme_lab() +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "bottom")
```

# Correlation

### Scatterplot
```{r}
ggplot(filter(gapminder, year == 2007) , aes(x = gdpPercap, y = lifeExp)) +
  geom_point(colour = '#fc6721') +
  scale_x_log10(labels = scales::dollar) +
   labs(title = "",
       subtitle = "Relationship between life expectancy and income, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = "GDP per capita ($)",
       y = "Age (years)") +
    theme_lab() +
  theme(panel.grid.major.x = element_blank())
```

### Bubble chart

```{r}
ggplot(filter(gapminder, year == 2007), aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(size = pop, fill = continent), shape = 21, colour = "white") +
  scale_x_log10(labels = scales::dollar) +
  scale_fill_brewer(palette = "Set2") +
  scale_size_continuous(range = c(1, 20)) +
  labs(title = "",
       subtitle = "Relationship between life expectancy and income, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = "GDP per capita ($)",
       y = "Age (years)" )  +
  guides(size = FALSE) +
   theme_lab() +
   theme(panel.grid.major.x = element_blank(),
        legend.position = "right", 
        legend.title = element_blank())
```

# Deviation

### Diverging bar chart
```{r}
df <- gapminder %>%
  filter(year == 2007 & continent == "Europe") %>%
  mutate(median = median(gdpPercap),
         diff = gdpPercap - median,
         type = ifelse(gdpPercap < median, "Below", "Above")) %>% 
  arrange(diff) %>% 
  mutate(country = factor(country, levels = country))

ggplot(df, aes(x = country, y = diff, label = country)) + 
  geom_col(aes(fill = type), width = 0.5, alpha = 0.8)  +
  scale_y_continuous(expand = c(0, 0), labels = scales::dollar) +
  scale_fill_manual(labels = c("Above median", "Below median"), 
                    values = c("Above" = "#31a354", "Below" = "#de2d26")) + 
  labs(title = "",
       subtitle = "GDP per capita, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = NULL,
       y = NULL,
       fill = NULL) + 
  coord_flip() +
  theme_lab() +
  theme(panel.grid.major.y = element_blank())
```

# Distribution

### Histogram

```{r}
ggplot(filter(gapminder, year == 2007), aes(x = lifeExp)) +
  geom_histogram(binwidth = 1, fill = '#fc6721', colour = "white" , alpha = 0.8) +
   scale_y_continuous(breaks = scales::pretty_breaks()) +
  labs(title = "",
       subtitle = "Life expectancy distribution, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = "Age (years)",
       y = "") +
  theme_lab() +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) 
  
  
```

### Density plot

```{r}
ggplot(filter(gapminder, year == 2007), aes(x = lifeExp)) + 
  geom_density(aes(fill = continent), size = 0.1, alpha = 0.5) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "",
       subtitle = "Life expectancy distribution, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = "Age (years)",
       y = "",
       fill = NULL) +
  theme_lab() +
  theme(panel.grid.major.x = element_blank())
```

### Boxplot

```{r}
ggplot(filter(gapminder, year == 2007), aes(x = continent, y = lifeExp, fill = continent)) + 
  geom_boxplot(colour = "#757575", alpha = 0.8) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "",
       subtitle = "Life expectancy distributions, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = "",
       y = "Age (years)") +
  theme_lab() +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "none")
```

### Violin plot

```{r}
ggplot(filter(gapminder, year == 2007), aes(x = continent, y = lifeExp, fill = continent)) +
  geom_violin(colour = "#757575", alpha = 0.8) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "",
       subtitle = "Life expectancy distribution, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = "",
       y = "Age (years)") +
  theme_lab() +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "none")
```

### Ridgeline plot

```{r}
library(ggridges)
df <- gapminder %>% filter(year == 2007 & continent != "Oceania")

ggplot(df, aes(x = lifeExp, y = fct_rev(continent), fill = continent)) +
  geom_density_ridges(colour = "#bdbdbd", size = 0.5, alpha = 0.5) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "",
       subtitle = "Life expectancy distribution, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = "Age (years)",
       y = "") +
  theme_lab() +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "none")
```

# Flow Magnitude

## Bar chart (vertical)

```{r}
df <- gapminder %>%
      filter(year == 2007) %>%
      group_by(continent) %>%
      summarise(median = median(gdpPercap))

ggplot(df, aes(x = continent, y = median, fill = continent)) + 
  geom_col(alpha = 0.8) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = scales::dollar, expand = c(0, 0)) +
  labs(title = "",
       subtitle = "Median GDP per capita by continent, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = NULL,
       y = "GDP per capita",
       fill = NULL) +
   theme_lab() +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "none")
```

## Grouped bar chart
```{r}
df <- gapminder %>% 
  filter(year > 1990) %>%
  group_by(year, continent) %>%
  summarise(totalpop = sum(as.double(pop)))

ggplot(df, aes(x = year, y = totalpop, group = continent, fill = continent)) + 
  geom_col(position = "dodge", colour = "#757575", size = 0.2, alpha = 0.8) +
  scale_x_continuous(breaks = seq(1992, 2007, 5), expand = c(0, 0)) +
  scale_y_continuous(labels = scales::comma, expand = c(0, 0)) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "",
       subtitle = "Total population by continent, 1990-2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = NULL,
       y = NULL,
       fill = NULL) +
  theme_lab() +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "bottom")
```

## Stacked bar chart

```{r}
df <- gapminder %>%
        filter(year > 1990) %>%
        group_by(year, continent) %>%
        summarise(totalpop = sum(as.double(pop)))

ggplot(df, aes(x = year, y = totalpop, fill = continent)) + 
  geom_col(colour = "white", size = 0.2, alpha = 0.8) +
   scale_x_continuous(breaks = seq(1992, 2007, 5), expand = c(0, 0)) +
   scale_y_continuous(labels = scales::comma, expand = c(0, 0)) +
   scale_fill_brewer(palette = "Set2") +
   guides(fill = guide_legend(reverse = T)) +
   labs(title = "",
       subtitle = "Total population by continent, 1990-2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = NULL,
       y = "Population",
       fill = NULL) +
   theme_lab() +
   theme(panel.grid.major.x = element_blank(),
        legend.position = "right")
```
# Part-to-whole

## 100% stacked bar chart


```{r}
df <- gapminder %>% 
  filter(year > 1990) %>%
  group_by(year, continent) %>%
  summarise(totalpop = sum(as.double(pop)))

ggplot(df, aes(x = year, y = totalpop, fill = continent)) + 
  geom_col(position = "fill", colour = "#757575", size = 0.2, alpha = 0.8) +
  scale_x_continuous(breaks = seq(1992, 2007, 5), expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent, expand = c(0, 0)) +
  scale_fill_brewer(palette = "Set2") +
  guides(fill = guide_legend(reverse = T)) +
  labs(title = "",
       subtitle = "Proportion of total population by continent, 1990-2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = NULL,
       y = NULL,
       fill = NULL) +
  theme_lab() +
  theme(panel.grid.major.x = element_blank(),
        legend.position = "right")
```

# Ranking

## Ordered bar chart (horizontal)


```{r}
df <- gapminder %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarise(median = median(gdpPercap))

ggplot(df, aes(reorder(continent, -median, sum), median)) +
  geom_col(fill = "#fc6721", alpha = 0.8) +
  scale_y_continuous(expand = c(0, 0), labels = scales::dollar) +
  coord_flip() +
  labs(title = "",
       subtitle = "Median GDP per capita by continent, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = NULL,
       y = "GDP per capita",
       fill = NULL) +
  theme_lab() +
  theme(panel.grid.major.y = element_blank())
```

## Lollipop chart

```{r}
df <- gapminder %>% 
  filter(year == 2007 & continent == "Europe") %>% 
  arrange(gdpPercap) %>% 
  mutate(country = factor(country, levels = country))

ggplot(df, aes(x = gdpPercap, y = country)) +
  geom_segment(aes(x = 0, xend = gdpPercap, y = country, yend = country), colour = "#f0f0f0") + 
  geom_point(colour = "#fc6721", size = 3, alpha = 0.8) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, max(df$gdpPercap) * 1.1),
                     labels = scales::dollar) +
  labs(title = "",
       subtitle = "GDP per capita in European countries, 2007",
       caption = "Source: Gapminder.org  |  @traffordDataLab",
       x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme_lab() +
  theme(panel.grid.major = element_blank(),
        axis.text.y = element_text(hjust = 0))

```



# Sale Report General

## Sales Report {.tabset}

### By Product

(tab content)

### By Region

(tab content)


# Heading {.tabset .tabset-dropdown}

## Dropdown 1

## Dropdown 2

## Dropdown 3 

## Dropdown 4