---
title: "R Notebook"
author: "Data Table Format"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
      toc: TRUE
      toc_float: TRUE
---

# Create the percentage in the datas

### Finding percentage in a sub-group using group_by and summarise
reference: https://stackoverflow.com/questions/29549731/finding-percentage-in-a-sub-group-using-group-by-and-summarise
```{r include=FALSE}
library(dplyr)
```

### Constract the data
```{r}
df <- data.frame(
month=c("Feb-14", "Feb-14", "Feb-14", "Mar-14", "Mar-14", 
        "Mar-14", "Apr-14", "Apr-14", "Apr-14", "May-14"),
type=c("bbb", "ccc", "aaa", "bbb", "ccc", "aaa", "bbb", "ccc", "aaa", "bbb"),
count=c(341, 527, 2674, 811, 1045, 4417, 1178, 1192, 4793, 916))

df %>%
  group_by(month) %>%
  mutate(per = paste0(round(count/sum(count)*100,2),'%')) %>%
  ungroup()
```
### data percentage of the data
```{r}
df %>%
    group_by(month) %>%
    mutate(countT= sum(count)) %>%
    group_by(type, add=TRUE) %>%
    mutate(per=paste0(round(100*count/countT,2),'%'))
```

# Calculate month on month change in R

reference: https://stackoverflow.com/questions/38055475/calculate-month-on-month-change-in-r
```{r}
dt<-data.frame(c_name=c("ABC","ABC","ABC","DEF","DEF","DEF"), 
               pmonth=c(201401,201402,201403,201401,201402,201403), 
               spend=c(100,150,230,110,190,300))

head(dt)
```

### data manpulation
```{r}
dt %>%
  group_by(c_name) %>%
  mutate(mom_change = c(NA, diff(spend)), 
         mom_per_change = (spend - lag(spend))/lag(spend))
```

### Calculate Month-On-Month Change In Website Traffic 
reference: https://analyticslog.com/blog/how-to-calculate-month-on-month-change-in-website-traffic-using-ggplot-in-r-studio
```{r}
gadata = data.frame(yearMonth=c(201801, 201802,201803,
                                201804,201805, 201806,
                                201807, 201808 ,201809), 
                    sessions=c(145,126,202,224,214,248,255,286,264))
gadata <- gadata %>%
  mutate(yearMonth = as.factor(yearMonth))

head(gadata)
```

```{r}
# new column added using a calculated field 

gadata <- gadata %>%
            mutate(mom_per_change = (sessions - lag(sessions))/lag(sessions))
head(gadata)
```

```{r}
ggplot(gadata, mapping=aes(x=yearMonth, 
  y=mom_per_change)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent)
```

# dplyr filter() with SQL-like %wildcard%
https://stackoverflow.com/questions/32829358/dplyr-filter-with-sql-like-wildcard
```{r}
library(dplyr)
library(data.table)
library(stringr)


foo <- data.frame(Company = c("company1", "foo", "test", "food"), 
                  Metric = rnorm(4, 10))

head(foo)
```

```{r}
foo %>% 
  filter(str_detect(Company,"foo"))
```
```{r}
foo %>% filter(str_detect(Company,"^f")) 
```

```{r}
foo %>%
  filter(Company %like% "foo")
```

# Identify and Remove Duplicate Data in R
https://www.datanovia.com/en/lessons/identify-and-remove-duplicate-data-in-r/
```{r}
suppressWarnings(library(tidyverse))


my_data = as_tibble(iris)
head(my_data,3)
```
```{r}
dim(my_data)
```

```{r}
# Remove duplicates based on Sepal.Width columna
Sepal.Width.Remove.duplicates <- my_data[!duplicated(my_data$Sepal.Width),]
dim(Sepal.Width.Remove.duplicates)
```

```{r}
# Remove duplicate rows based on Sepal.Length 
my_data %>% distinct(Sepal.Length, .keep_all = TRUE)  %>% head()
```
```{r}
# Remove duplicated rows based on 
# Sepal.Length and Petal.width

my_data %>% 
  distinct(Sepal.Length, Petal.Width, .keep_all = TRUE) %>% head()
```




# How to suppress warning messages when loading a library?
https://stackoverflow.com/questions/18931006/how-to-suppress-warning-messages-when-loading-a-library

```{r}
shhh <- suppressPackageStartupMessages 

shhh(library(tidyverse))
```

# Subset of rows containing NA (missing) values in a chosen column of a data frame

https://stackoverflow.com/questions/7980622/subset-of-rows-containing-na-missing-values-in-a-chosen-column-of-a-data-frame
```{r}
new_data <- data %>% filter_all(any_vars(is.na(.))) 

```

# Replace NA with Zero in dplyr without using list()
https://stackoverflow.com/questions/49947592/replace-na-with-zero-in-dplyr-without-using-list
```{r}
library(dplyr)
df <- tibble(x = c(1, 2, NA), 
             y = c("a", NA, "b"), 
             z = list(1:5, NULL, 10:20))
df
```
```{r}
df %>% 
    replace_na(list(x = 0, y = "unknown"))
```

```{r}
df$x[is.na(df$x)] <- 0
```

